An event-sourced and decentralized database with IPLD and materialized views

### 2019-09-24 draft.tex
update: add ref and broadcaster info
Signed-off-by: Carson Farmer <carson.farmer@gmail.com>

To make Threads as easy to adopt and use as possible, we've designed the internal event store to take advantage of ideas from several existing CQRS and ES systems, as well as ideas and designs from Flux \cite{facebookFluxInDepthOverview2019} and Redux\footnote{Redux builds on concepts from CRQS and ES itself, and is arguably an implementation of the Flux application architecture.} \cite{reduxMotivation} and domain driven design \cite{evansDomaindrivenDesignTackling2004a} (DDD)\footnote{\url{https://dddcommunity.org}}. In this section, we outline the internal structure of a local Threads ``client'' at a relatively high-level, and outline how it can be used to build intuitive developer-facing APIs to make adopting and using Threads ``the right choice'' for a wide range of developers.

@@

The Dispatcher, combined with Actions and Creators provide a useful interface for the write side of the Threads CQRS framework. Like many CQRS-based systems, the read side revolves around the concept of views (or Stores). As Actions are dispatched on the write side, they are reacted to on the read side via Reducers, causing updates to Read/View Models, which provide interfaces for queries and accessing state. In practice, a ReadStore is defined by a Reducer function and a set of query Resolver functions. This is very similar to a Redux Store, though it is possible to have \emph{multiple} Stores as in the more general Flux pattern, which is a departure of the ``single source of truth'' in Redux.

\begin{definition}
  (Reducer). Reducers are ``pure'' functions (where possible)\footnote{In practice they aren't really pure, because we are mutating (persistent) state here}, and are written to modify a ReadStore's state based on incoming Actions.  
\end{definition}

@@

Threads also define \emph{ViewStores}, which are a special kind of ReadStore, akin to a Redux Store. In practice, they are smaller, read-only (immutable) models, that are built on-the-fly, and defined by a pure Reducer function. Textile Threads use ViewStores to define the ACL for a given Thread.

@@

A database built on Threads would map database operations into Actions. Here, the \emph{domain} is the databse. These Actions would then mutate a ReadModel via its corresponding Reducer function, as per above. The Store could be persisted or kept in-memory (as a ViewModel), and would be directly queried via corresponding database-style Resolver metods. Because Textile uses ES patterns, any delete-type operation would simply lead to an internal \emph{tombstone} Action, marking the database item for removal from the ReadStore.

@@

Other database abstractions include a no-sql style document store for storing and indexing arbitrary structs and/or JSON documents. The interface for such as store, again built using a materialized \texttt{ReadStore}, might look like Figure \ref{ex:DocStore}, where \texttt{Indexable} could be satisfied by any structure with a \texttt{Key} field and \texttt{Query} might be taken from the \texttt{go-datastore} interface library\footnote{\url{https://github.com/ipfs/go-datastore}} or similar.

@@

At an abstract level, the Textile protocol provides a highly distributed framework for building shared, offline first, Stores that are fault tolerant\footnote{When using an ACID compliant backing store for example.}, eventually consistent, and scalable. Any internal implementation details of a compliant Threads ``client'' may use any number of well-established design patterns from the CQRS and ES (and related) literature to ``extend'' the Threads protocol with additional features and controls. Indeed, by designing our system around Actions, a Dispartcher, and generic ReadStores, we aim to make easy to extend Threads in many different ways. Some examples of extensions that are included by default in Textile's Threads implementation are outlined in this section to provide some indication of the extensibility this design affords.

>>>>

To make Threads as easy to adopt and use as possible, we've designed the internal event store to take advantage of ideas from several existing CQRS and ES systems (e.g., \cite{ereminReduxInspiredBackend2019}), as well as ideas and designs from Flux \cite{facebookFluxInDepthOverview2019}, Redux\footnote{Redux builds on concepts from CRQS and ES itself, and is arguably an implementation of the Flux application architecture.} \cite{reduxMotivation} and domain driven design \cite{evansDomaindrivenDesignTackling2004a} (DDD)\footnote{\url{https://dddcommunity.org}}. In this section, we outline the internal structure of a local Threads ``client'' at a relatively high-level, and outline how it can be used to build intuitive developer-facing APIs to make adopting and using Threads ``the right choice'' for a wide range of developers.

@@

The Dispatcher, combined with Actions and Creators provide a useful interface for the write side of the Threads CQRS framework. Like many CQRS-based systems, the read side revolves around the concept of views (or Stores). As Actions are dispatched on the write side, they are reacted to on the read side via Reducers, causing updates to Read/View Models, which provide interfaces for queries and accessing state. Read Stores are also responsible for notifying downstream consumers (application logic) of changes to their state via \emph{broadcast channels}. In practice, a Store is defined by a Reducer function, a set of query Resolver functions \cite{ereminReduxInspiredBackend2019}, and an Event Bus (or Broadcast channel). This similar in some respects to a Redux Store, though it is possible to have \emph{multiple} Stores as in the more general Flux pattern.

\begin{definition}
  (Reducer). Reducers are ``pure'' functions (where possible)\footnote{In practice they aren't really pure, because we are mutating (persistent) state here}, and are written to modify a Store's state based on incoming Actions.  
\end{definition}

@@

\begin{definition}
  (Store). Stores in Threads provide flexible APIs for persisting and accessing state. They are defined by a Reducer function for handling incoming Actions, and a set of query Resolver functions for handling queries. They are additionally responsible for notifying downstream comsumers of updates to their state via a Broadcaster.
\end{definition}

@@

A database built on Threads would map database operations into Actions. Here, the \emph{domain} is the databse. These Actions would then mutate a ReadModel via its corresponding Reducer function, as per above. The Store could be persisted or kept in-memory (as a ViewModel), and would be directly queried via corresponding database-style Resolver metods. Because Textile uses ES patterns, any delete-type operation would simply lead to an internal \emph{tombstone} Action, marking the database item for removal from the Store.

@@

Other database abstractions include a no-sql style document store for storing and indexing arbitrary structs and/or JSON documents. The interface for such as store, again built using a materialized Store, might look like Figure \ref{ex:DocStore}, where \texttt{Indexable} could be satisfied by any structure with a \texttt{Key} field and \texttt{Query} might be taken from the \texttt{go-datastore} interface library\footnote{\url{https://github.com/ipfs/go-datastore}} or similar.

@@

At an abstract level, the Textile protocol provides a highly distributed framework for building shared, offline first, Stores that are fault tolerant\footnote{When using an ACID compliant backing store for example.}, eventually consistent, and scalable. Any internal implementation details of a compliant Threads ``client'' may use any number of well-established design patterns from the CQRS and ES (and related) literature to ``extend'' the Threads protocol with additional features and controls. Indeed, by designing our system around Actions, a Dispartcher, and generic Stores, we aim to make easy to extend Threads in many different ways. Some examples of extensions that are included by default in Textile's Threads implementation are outlined in this section to provide some indication of the extensibility this design affords.
